Model-based pretreatment for regression with correlated variables.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@489 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage[table]{xcolor}
 \graphicspath{{figures/}}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
\title{CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry}

@@

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
We note the linear regression model:
\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}). As shown in section \ref{sectionOLS}, estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\hat{\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model  : the more covariates you have the greater variance you get.
	\item The correlations within the covariates : strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

@@

In the following we note classical norms: $\parallel\beta\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\beta \parallel_1=\sum_{i=1}^p|\beta_i| $.

	Ridge regression\cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:
	\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_2^2\leq k
	\end{equation}
	But Ridge regression do not aim at selecting covariates because coefficients tends to 0 but don't reach 0.
	So it gives difficult interpretations for large values of $p$.

	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates. And it  really selects covariates with coefficients set exactly to 0.
	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
	 \\

	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}. 
%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
%	\begin{equation}
%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
%\end{equation}		
	Elastic net can be written:
	\begin{equation}
		\hat{\beta}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel Y-X\beta \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\beta\parallel_1+\alpha\parallel\beta\parallel_2^2\leq t \textrm{ for some } t
	\end{equation}
	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
	%It seems to give good predictions. 
	But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model. Grouping effect is the fact that correlated covariates get similar coefficient and are selected together.
	\\

@@

\section{Model to select decorrelated covariates}
\subsection{The decorrelation problem in regression}\label{sectionOLS}
For a model defined by 
	\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
	\begin{equation}
		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
	\end{equation}
	With variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1} \label{eqOLS}
	\end{equation}
	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
	\begin{eqnarray}
		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
	\end{eqnarray}	 
	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
	Another problem is that matricial inversion requires $n\geq p$. When it is not the case, a frequently used alternative is the Moore-Penrose generalized inverse \cite{katsikis2008fast}. In the followings, matrices inverses when $n<p$ will refer to this generalized inverse. Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
\subsection{Modelisation of the correlations}
Let now $X \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$. Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $p_1=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression. Thus we have $p_2=| I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage[table]{xcolor}
 \graphicspath{{figures/}}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
\title{Model-based pretreatment for regression with correlated variables.}

@@

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
	\end{equation}
	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}). %As shown in section \ref{sectionOLS}, 
	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

@@

In the following we note classical norms: $\parallel\boldsymbol{\beta}\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\boldsymbol{\beta} \parallel_1=\sum_{i=1}^p|\beta_i| $.

	Ridge regression\cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:
	\begin{equation}
		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel \boldsymbol{\beta} \parallel_2^2\leq k
	\end{equation}
	But Ridge regression do not aim at selecting covariates because coefficients tends to 0 but don't reach 0.
	So it gives difficult interpretations for large values of $p$.

	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates. And it  really selects covariates with coefficients set exactly to 0.
	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
	 \\

	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}. 
%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
%	\begin{equation}
%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
%\end{equation}		
	Elastic net can be written:
	\begin{equation}
		\boldsymbol{\hat{\beta}}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta} \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\boldsymbol{\beta}\parallel_1+\alpha\parallel\boldsymbol{\beta}\parallel_2^2\leq t \textrm{ for some } t
	\end{equation}
	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
	%It seems to give good predictions. 
	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
	\\

@@

\section{Model to select decorrelated covariates}
\subsection{A classical problem: correlations in regression}\label{sectionOLS}
For a model defined by 
	\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
	\begin{equation}
		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
	\end{equation}
	With variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1} \label{eqOLS}
	\end{equation}
	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
	\begin{eqnarray}
		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
	\end{eqnarray}	 
	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
	Another problem is that matricial inversion requires $n\geq p$. When it is not the case, a frequently used alternative is the Moore-Penrose generalized inverse \cite{katsikis2008fast}. In the followings, matrices inverses when $n<p$ will refer to this generalized inverse. Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
\subsection{Modelisation of the correlations}
Let now $X \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$. Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $p_1=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression. Thus we have $p_2=| I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.
